Optimal (Multiway) Spatial Joins

### 2024-05-30 msj.tex
yufei

\extraspacing {\bf Algorithm for Problem $\bm{\mathscr{A}}$.}

>>>>

\extraspacing {\bf Algorithm for Problem $\bm{\mathscr{A}}$.} The {\em interval tree} \cite{bcko08} stores a set $S$ of $N$ intervals in $O(N)$ space such that, given any real value $q$, the intervals of $S$ containing $q$ can be found in $O(\log N + K)$ time, where $K$ is the number of intervals reported. It can also be used to detect whether $S$ has at least one interval containing $q$ in $O(\log N)$ time. Each interval insertion or deletion on $S$ can be supported in $O(\log N)$ time.

Let us first consider the detection version of Problem $\mathscr{A}$. Each point $p = (x, y)$ of $P$ is said to {\em define} the y-coordinate $y$, and each rectangle $r = [x_1, x_2] \times [y_1, y_2]$ is said to {\em define} two y-coordinates $y_1$ and $y_2$. We start by sorting the set of y-coordinates defined by all the points and rectangles.

Next, scan (conceptually) a horizontal line $\ell$ from $y = -\infty$ to $y = \infty$. At all times, maintain the set $R_\ell$ of rectangles in $R$ that intersect with $\ell$. Let $S_\ell$ be the set of x-ranges of the rectangles in $R_\ell$; we store $S_\ell$ in an interval tree $\T$. Specifically, when $\ell$ hits the bottom (resp., top) edge of a rectangle $r = [x_1, x_2] \times [y_1, y_2]$ of $R$, we insert (resp., delete) $[x_1, x_2]$ into (resp., from) $\T$, which can be done in $O(\log n)$ time. When $\ell$ hits a point $p = (x, y)$ of $P$, search $\T$ to determine if any interval in $S_\ell$ contains the value $x$. If so, point $p$ is covered by at least one rectangle in $R$; otherwise, it is not. The overall running time is $O(n \log n)$.

The algorithm for the reporting version of Problem $\mathscr{A}$ is similar. The only difference is that, when $\ell$ hits a point $p = (x, y)$ of $P$, we use $\T$ to report all the intervals in $S_\ell$ that contain $x$; the cost is $O(n \log n + K_p)$, where $K_p$ is the number of such intervals. Every interval corresponds to a rectangle in $R$ that contains $p$. The total running time is $O(n \log n + \sum_p K_p) = O(n \log n + \out)$.
